------------------------------------------------------------------------

---
title: "Agreement Initial Analysis"
author: "Chloe Mirzayi"
format: html
editor: visual
---

# Download spreadsheets
NOTE FOR USE: Link to private google sheet removed. Instead load from data/aim 1/Aim 1 Agreement.xlsx.
```{r, eval=FALSE}
library(gsheet)
library(reshape2)
library(tidyr)

url <- ""
studies <- gsheet2tbl(url)

alldata <- lapply(studies$url, gsheet2tbl)

```

# Study Agreement and Reliability

## Agreement

```{r}
library(dplyr)
library(irr)
agreeStudy <- function(ds){
  ds <- ds %>% mutate_all(.,toupper)
  agree(ds[,2:ncol(ds)])$value %>% return()
}
```

```{r}
testblah <- lapply(data, agreeStudy) %>% melt() %>% select(1)
  lapply(alldata, agreeStudy) %>% melt() %>% select(1) %>% write.csv(file="studyagree.csv")
```

## Reliability

```{r}
kappaStudy <- function(ds){
  ds <- ds %>% mutate_all(.,toupper)
  kappam.fleiss(ds[,2:ncol(ds)], detail = TRUE) %>% return()
}
```

```{r}
lapply(alldata, kappaStudy)
kappam.fleiss(alldata[[17]][,2:3])
```

# Item Agreement

```{r}
library(matrixStats)
iagree <- function(item){
  
  rowagree <- vector(mode="list", length=length(data))
  #Calculate agreement for that row in each dataset
  for (i in 1:length(data)){
  ds <- data[[i]] %>% mutate_all(~replace(., is.na(.), "MISSING")) %>% mutate_all(.,toupper)
  row <- ds[item,]
  denom <- 6
  
  x <- 0
  if (row[,2]=="YES") {x <- x+1}
  if (row[,3]=="YES") {x <- x+1}
  if (row[,4]=="YES") {x <- x+1}
  num1 <- x*(x-1)
  
  x <- 0
  if (row[,2]=="NOT PRESENT") {x <- x+1}
  if (row[,3]=="NOT PRESENT") {x <- x+1}
  if (row[,4]=="NOT PRESENT") {x <- x+1}
  num2 <- x*(x-1)
  
  x <- 0
  if (row[,2]=="NO") {x <- x+1}
  if (row[,3]=="NO") {x <- x+1}
  if (row[,4]=="NO") {x <- x+1}
  num3 <- x*(x-1)
  
  rowagree[[i]] = num1/denom + num2/denom + num3/denom
  }

  agree <- rowagree %>% melt() %>% select(1)
  meanagree <- colMeans(agree)
  sdagree <- sd(agree[[1]])
  data.frame(item=ds[item,1], agreement=meanagree, sd=sdagree) %>% return()
}
```

```{r}
library(reshape2)

data <- alldata[-c(16,17,29)]
listagree <- lapply(1:69, iagree)

library(tidyr)
listagree %>% melt() %>% pivot_wider(id_cols=Number, names_from=variable, values_from=value) %>% write.csv(file="itemagree.csv")

```

# Study Agreement

# Were Not Presents more likely for animal studies?

```{r}
library(stringr)
studysum <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1BsAQU-zDCyt9kt4Ej2433o-46cCM2WcJl6zHPADFHrQ/edit#gid=879139569")

head(studysum)
numsbytype <- studysum %>% mutate(studynum = str_sub(study_id,-2)) %>% group_by(type) %>% summarize(studynum=list(unique(as.integer(studynum))))

countstudies <- numsbytype$studynum %>% unlist
#countstudies <- countstudies[-length(countstudies)] activate for cross-sectional
alldata[countstudies] %>% unlist() %>% table(.,.>length(countstudies))
```

# Agreement by Study Type

## Overall

```{r}
library(reshape2)
library(tidyr)
data <- alldata[-c(16,17,29)]
allagree <- lapply(1:69, iagree)
all <- allagree %>% melt() %>% pivot_wider(id_cols=Number, names_from=variable, values_from=value)

mean(all$agreement)
sd(all$agreement)

all %>% write.csv(file="allstudiesagree.csv")
```

## RCT

```{r}
countstudies <- numsbytype$studynum[numsbytype$type=="RCT"] %>% unlist
countstudies

data <- NA
data<- alldata[countstudies]
data <- data[-c(4)]
rctagree <- lapply(1:69, iagree)

rct <- rctagree %>% melt() %>% pivot_wider(id_cols=Number, names_from=variable, values_from=value)

mean(rct$agreement)
sd(rct$agreement)

rct %>% write.csv(file="rctagree.csv")
```

## Cross-Sectional

```{r}
countstudies <- numsbytype$studynum[numsbytype$type=="cross-sectional"] %>% unlist
countstudies

data <- NA
data<- alldata[countstudies]
data <- data[-c(10)]
csagree <- lapply(1:69, iagree)

cs <- csagree %>% melt() %>% pivot_wider(id_cols=Number, names_from=variable, values_from=value)

mean(cs$agreement)
sd(cs$agreement)

cs %>% write.csv(file="csagree.csv")
```

## Longitudinal

```{r}
countstudies <- numsbytype$studynum[numsbytype$type=="longitudinal"] %>% unlist
countstudies

data <- NA
data<- alldata[countstudies]
longagree <- lapply(1:69, iagree)

long <- longagree %>% melt() %>% pivot_wider(id_cols=Number, names_from=variable, values_from=value)

mean(long$agreement)
sd(long$agreement)

long %>% write.csv(file="longagree.csv")
```

## Animal

```{r}

countstudies <- numsbytype$studynum[numsbytype$type=="animal"] %>% unlist
countstudies

data <- NA
data<- alldata[countstudies]
data <- data[-c(4)]
anagree <- lapply(1:69, iagree)

an <- anagree %>% melt() %>% pivot_wider(id_cols=Number, names_from=variable, values_from=value)

mean(an$agreement)
sd(an$agreement)

an %>% write.csv(file="anagree.csv")
```

```{r}
studyagreement <- function(study){
ds <- as.data.frame(alldata[study]) %>% mutate_all(~replace(., is.na(.), "MISSING")) %>% mutate_all(.,toupper)
rowagree <- vector(mode="list", length=length(data))
for (i in 1:69){
  row <- ds[i,]
  denom <- 6
  
  x <- 0
  if (row[,2]=="YES") {x <- x+1}
  if (row[,3]=="YES") {x <- x+1}
  if (row[,4]=="YES") {x <- x+1}
  num1 <- x*(x-1)
  
  x <- 0
  if (row[,2]=="NOT PRESENT") {x <- x+1}
  if (row[,3]=="NOT PRESENT") {x <- x+1}
  if (row[,4]=="NOT PRESENT") {x <- x+1}
  num2 <- x*(x-1)
  
  x <- 0
  if (row[,2]=="NO") {x <- x+1}
  if (row[,3]=="NO") {x <- x+1}
  if (row[,4]=="NO") {x <- x+1}
  num3 <- x*(x-1)
  
  rowagree[[i]] = num1/denom + num2/denom + num3/denom
}

  agree <- rowagree %>% melt() %>% select(1)
  meanagree <- colMeans(agree)
  sdagree <- sd(agree[[1]])
  data.frame(agreement=meanagree, sd=sdagree) %>% return()
}

threeevals <- c(1:15, 18:28)
lapply(threeevals, studyagreement) %>% melt() %>% pivot_wider(id_cols=L1, names_from=variable, values_from=value) %>% write.csv(file="studyagree.csv")
```

```{r}
studyagreement <- function(i){
  data <- alldata[i]
  studyagree <- lapply(1:69, iagree)
  
  study <- studyagree %>% melt() %>% pivot_wider(id_cols=Number, names_from=variable, values_from=value)
  
  data.frame(mean = mean(study$agreement), sd = sd(study$agreement)) %>% return()
}

listagree <- lapply(1:28, studyagreement)

listagree %>%  melt() %>% pivot_wider(id_cols=L1, names_from=variable, values_from=value)
```

# Multilevel model of agreement

```{r}
# Make a long data frame
data <- alldata[-c(16,17,29)]

colnames <- c("Number", "E1", "E2", "E3")
datalong <- lapply(alldata, setNames, colnames) %>% melt()
datalong <- datalong[datalong$L1 != 16 & datalong$L1 != 17 & datalong$L1 != 29,]

# Calculate agreement for each item in the data frame

rowagree <- vector(mode="list", length=nrow(datalong))
  #Calculate agreement for each row
  for (i in 1:nrow(datalong)){
  row <- datalong[i,] %>%  mutate_all(~replace(., is.na(.), "MISSING")) %>% mutate_all(.,toupper)
  denom <- 6
  
  x <- 0
  if (row[,2]=="YES") {x <- x+1}
  if (row[,3]=="YES") {x <- x+1}
  if (row[,4]=="YES") {x <- x+1}
  num1 <- x*(x-1)
  
  x <- 0
  if (row[,2]=="NOT PRESENT") {x <- x+1}
  if (row[,3]=="NOT PRESENT") {x <- x+1}
  if (row[,4]=="NOT PRESENT") {x <- x+1}
  num2 <- x*(x-1)
  
  x <- 0
  if (row[,2]=="NO") {x <- x+1}
  if (row[,3]=="NO") {x <- x+1}
  if (row[,4]=="NO") {x <- x+1}
  num3 <- x*(x-1)
  
  rowagree[[i]] = num1/denom + num2/denom + num3/denom
  }

agreelong <- rowagree %>% melt() %>% select(1)

agreedatalong <- cbind(datalong, agreelong)
```

Label studies

countstudies

```{r}
countstudies <- numsbytype$studynum[numsbytype$type=="RCT"] %>% unlist
agreedatalong$studytype[agreedatalong$L1 %in% countstudies] <- "RCT"

countstudies <- numsbytype$studynum[numsbytype$type=="cross-sectional"] %>% unlist
agreedatalong$studytype[agreedatalong$L1 %in% countstudies] <- "cross-sectional"

countstudies <- numsbytype$studynum[numsbytype$type=="longitudinal"] %>% unlist
agreedatalong$studytype[agreedatalong$L1 %in% countstudies] <- "longitudinal"

countstudies <- numsbytype$studynum[numsbytype$type=="animal"] %>% unlist
agreedatalong$studytype[agreedatalong$L1 %in% countstudies] <- "animal"
table(agreedatalong$studytype)
```

Ok now that we have agreement let's try a multilevel model

```{r}

library(lme4)
library(lmerTest)
lmer(value ~  studytype + (1|L1), data=agreedatalong) %>% summary()
lm(value ~ studytype, data=agreedatalong) %>% summary()

```
